Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-06-16 paper_enahpc14/Sec2:COSMO-ART.tex
updates paper according to page proofs

COSMO-ART  is  fully  online-coupled,  and  allows  for  feedbacks  of
aerosols  on  temperature,  radiation  and cloud  condensation  nuclei
(CCN).   Analytical  description  of  modules  incorporated  for  this
purpose  exists  in  \cite{Vogel-2009, Bangert-2011}.   The  radiation
scheme used  within the  model to calculate  the vertical  profiles of
shortwave and longwave  radiative fluxes is GRAALS \cite{Ritter-1992}.
In order to account for  the interaction of aerosol particles with the
cloud microphysics and radiation,  COSMO-ART uses the two-moment cloud
microphysics  scheme  of Seifert  and  Beheng \cite{Seifert-2006}  and
comprehensive parameterizations for cloud condensation and ice crystal
nucleation  \cite{Bangert-2011, Bangert-2012}.   The  system of  stiff
ordinary differential equations described by chemical reactions in the
aqueous-phase together with the transfer reactions is solved using the
Kinetic PreProcessor (KPP, \cite{Damian-2002}).

Further model description details  can be found in \cite{Bangert-2012,
  Knote-2011, Knote-2013}.

@@

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for  April 13$^{th}$ 2010, near the equinox,
in  order  to  approximately have  a  half  day  of sun  exposure  and
therefore  ensure a typical  activation of  the chemistry  cycle.  The
calculation  domain  corresponds to  the  CORDEX-EU-44  domain and  is
covered  by  a  grid  of  $222\times 216$  points  with  a  horizontal
resolution  of $0.22\,^{\circ}$,  i.e., 50~km  in both  directions and
40~vertical  layers.   These  24-hour  forecast  simulations  are  not
preceded  by a spin-up  phase to  build up  the simulated  gaseous and
aerosol   concentrations.   This  condition   means  that   the  model
initialization needs  a period of adjustment to  reach its equilibrium
state  and minimise  the effect  of initial  conditions for  gases and
aerosols on the model predictions.

The meteorological initial and boundary conditions are obtained by the
the ECMWF  global spectral model IFS  with an update  frequency of 3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at  6h temporal resolution.   The model  setup incorporates  34~2D and
45~3D  fields to  be written  out  every hour  and is  devoid of  data
assimilation methods.

The   considered  \cosmoart   model  system   is  configured   with  a
semi-Lagrangian   horizontal    advection   sche\-me   with   tricubic
interpolation  and selective filling  diffusion option  in combination
with   the   dynamical    core   using   Runge-Kutta   time   stepping
\cite{COSMO-PartI-2011}.    It   also  makes   use   of  the   Kinetic
PreProcessor solver  (KPP) for  the solution of  atmospheric chemistry
ordinary  differential equations  \cite{Damian-2002}.   Concerning the
modelling  of  wet  deposition  in  aerosols, the  baseline  has  only
indirect  cloud feedbacks  but  does not  include in-cloud  scavenging
(rainout) and below-cloud  scavenging (washout) yet.  Amongst physical
parameterisations,   precipitation  formation   is   performed  by   a
two-moment cloud microphysics.

>>>>

COSMO-ART  is  fully  online-coupled,  and  allows  for  feedbacks  of
aerosols  on  temperature,  radiation  and cloud  condensation  nuclei
(CCN).   Analytical  description  of  modules  incorporated  for  this
purpose  exists  in  \cite{Vogel-2009, Bangert-2011}.   The  radiation
scheme used  within the  model to calculate  the vertical  profiles of
shortwave and longwave  radiative fluxes is GRAALS \cite{Ritter-1992}.
In order to account for  the interaction of aerosol particles with the
cloud microphysics and radiation,  COSMO-ART uses the two-moment cloud
microphysics  scheme  of Seifert  and  Beheng \cite{Seifert-2006}  and
comprehensive parameterizations for cloud condensation and ice crystal
nucleation  \cite{Bangert-2011, Bangert-2012}.   The  system of  stiff
ordinary differential equations described by chemical reactions in the
aqueous-phase together with the transfer reactions is solved using the
Kinetic   PreProcessor   (KPP,   \cite{Damian-2002}).  Further   model
description  details can be  found in  \cite{Bangert-2012, Knote-2011,
  Knote-2013}.

@@

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for  April 13$^{th}$ 2010, near the equinox,
in  order  to  approximately have  a  half  day  of sun  exposure  and
therefore  ensure a typical  activation of  the chemistry  cycle.  The
calculation  domain  corresponds to  the  CORDEX-EU-44  domain and  is
covered  by  a  grid  of  $222\times 216$  points  with  a  horizontal
resolution  of $0.22\,^{\circ}$,  i.e., 50~km  in both  directions and
40~vertical  layers.   These  24-hour  forecast  simulations  are  not
preceded  by a spin-up  phase to  build up  the simulated  gaseous and
aerosol   concentrations.   This  condition   means  that   the  model
initialization needs  a period of adjustment to  reach its equilibrium
state  and minimize  the effect  of initial  conditions for  gases and
aerosols on the model predictions.

The meteorological initial and boundary conditions are obtained by the
ECMWF  global spectral  model  IFS  with an  update  frequency of  3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at  6h temporal resolution.   The model  setup incorporates  34~2D and
45~3D  fields to  be written  out  every hour  and is  devoid of  data
assimilation methods.

The   considered  \cosmoart   model  system   is  configured   with  a
semi-Lagrangian   horizontal    advection   sche\-me   with   tricubic
interpolation  and selective filling  diffusion option  in combination
with   the   dynamical    core   using   Runge-Kutta   time   stepping
\cite{COSMO-PartI-2011}.    It   also  makes   use   of  the   Kinetic
PreProcessor solver  (KPP) for  the solution of  atmospheric chemistry
ordinary  differential equations  \cite{Damian-2002}.   Concerning the
modeling  of  wet  deposition  in  aerosols, the  baseline  has  only
indirect  cloud feedbacks  but  does not  include in-cloud  scavenging
(rainout) and below-cloud  scavenging (washout) yet.  Amongst physical
parameterizations,   precipitation  formation   is   performed  by   a
two-moment cloud microphysics.

### 2014-06-16 paper_enahpc14/Sec4:Experimental-results.tex
updates paper according to page proofs

To  gain  more  insights   about  the  power-performance  behavior  of
\cosmoart,  we  obtained power-per\-for\-man\-ce  traces  to allow  us
correlate  the type  of  tasks  being executed  with  the total  power
consumption  along  the   time  (see  Figure~\ref{fig:6}).   For  this
experiment we  limit the simulation  time to 1 hour  because: \emph{i})
the  pattern of  the computation  performed by  \cosmoart  is repeated
hourly and, \emph{ii}) the weight  of a 1-day simulation trace can not
be easily handled using our  machines and memory available (each trace
file was about 45\,GB). Thus, we shrink the simulations
to 1 hour during midday (from 11h  to 12h) and during
midnight  (from 23h  to  24h), in  which  \cosmoart perform  different
internal computations.  At  the same time we leverage  the two already
stated MPI policies: \emph{aggressive} and \emph{degraded}.

>>>>

To  gain  more  insights   about  the  power-performance  behavior  of
\cosmoart, we  obtained power-per\-for\-man\-ce traces to  allow us to
correlate  the type  of  tasks  being executed  with  the total  power
consumption  along  the   time  (see  Figure~\ref{fig:6}).   For  this
experiment we limit  the simulation time to 1  hour because: \emph{i})
the  pattern of  the computation  performed by  \cosmoart  is repeated
hourly and, \emph{ii}) the weight  of a 1-day simulation trace can not
be easily handled using our  machines and memory available (each trace
file was  about 45\,GB).   Thus, we shrink  the simulations to  1 hour
during midday (from 11h to 12h) and during midnight (from 23h to 24h),
in which  \cosmoart perform  different internal computations.   At the
same  time   we  leverage  the   two  already  stated   MPI  policies:
\emph{aggressive} and \emph{degraded}.
